---
title: "BikeSharingDemandChallenge"
author: "Chinmay Abhyankar"
output:
  github_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#read train data 
```{r warning=FALSE}
library(ggplot2);
library(stringi);

train = read.csv("train.csv");

```

#read test data 
```{r}
test = read.csv("test.csv");

```
#combine train and test data to see the distribution of independent variables together
```{r}
#test data has less attributes than train data
#first add all the missing attributes in test data
test$casual = 0
test$registered = 0
test$count = 0

datacombined = rbind(train,test)
```
#view the structure of data
```{r}
str(datacombined)
```
#missing values in data
```{r}
summary(is.na(datacombined))
```
#distribution of numerical values
```{r}
par(mfrow = c(4,2))
par(mar = rep(2, 4))
qplot(season, data = datacombined,binwidth = .1)
qplot(weather, data = datacombined,binwidth = .1)
qplot(humidity, data = datacombined,binwidth = .1)
qplot(holiday, data = datacombined,binwidth = .1)
qplot(workingday, data = datacombined,binwidth = .1)
qplot(temp, data = datacombined,binwidth = .1)
qplot(atemp, data = datacombined,binwidth = .1)
qplot(windspeed, data = datacombined,binwidth = .1)

```
#some inferences from the above plot
```{r}
#season has 4 categories of almost equal distribution
#weather has most of the data in 1st category
#less entries in the category 1 for holidays
#most entries in category 1 for working day attribute
#temp and atemp are naturally distributed
```

#converting discrete variables into factors

```{r}
names = c("season", "holiday","weather","workingday")
datacombined[,names] =lapply(datacombined[,names],FUN = as.factor)
str(datacombined)
```

#My hypothesis:counts should be more in peak hours
#lets verify my hypothesis

#we would use substring so first let us check the lenghth of datetime attribute
```{r}
stri_length(datacombined[1,]$datetime) #use stringi package
```
#use substr fuction

```{r}
datacombined$hour = substr(datacombined$datetime, 12, 13)
datacombined$datetime = as.factor(datacombined$datetime)
```

#I will analyze train only as test data does not have anything in count  
#therfore lets seperate train and test data
```{r}

train = datacombined[1:10886,]
test = datacombined[10887:17379,]
```


#analysis of count vs hour for training dataset
```{r}
qplot(hour, count, data = train, geom = "boxplot")
```
#Validation of my hypothesis:
#High number of counts between hours 7-9 and 17-19
#Average number of counts for hours 10-16
#Low number of counts between hours 0-6 and 20-24

#Now let us do same analysis for registered and casual users
```{r}
qplot(hour, registered, data = train, geom = "boxplot")
qplot(hour, casual, data = train, geom = "boxplot")
```
#My inference:Pattern for registered user is same as that of count
#whereas pattern for casual is little bit different.
#So we can conclude that hour is a significant factor for prediction

#let us analyze count on log scale to see if we can find some pattern
#also it will treat the outlier well 
```{r}
qplot(hour, log(count), data = train, geom = "boxplot")

```








